Urban Heat Risk Mapping Using Multiple Point Patterns in Houston, Texas

### 2016-08-16 jrssc-mpp.tex
Finished prior justification, began clarification of MCMC algorithm

A potential solution to avoid realigning the census data to the grid cells is to let the census block groups equal the $\mathcal{G}_k$ denoted in \eqref{lambda} rather than the grid cells. However, the heat information used in this research (described in detail in Section \ref{sec:temp_data}) is provided at a 1-km$^2$ resolution and these grid cells were chosen to align exactly with the available heat data. Using the block groups in place of the grid cells would necessitate aligning the heat information to the census block group areas. Because heat is our primary variable of interest we prefer realigning the demographic information to the grid associated with the heat data as opposed to the alternative. Therefore, we proceed having aligned the population data as described.

@@

\subsection{Hyperprior Layer}
At the hyperior prior layer of the model, we outline our prior assumptions for the model 
parameters in Sections \ref{likelayer} and \ref{process}. Because $\boldsymbol{\alpha}_p = 
(\alpha_p(0),\dots,\alpha_p(113))'$ are probabilities where $\sum_a \alpha_p(a) = 1$, 
we assume $$\boldsymbol{\alpha}_{p} \sim \text{Dir}(\mathbf{1}_{113})$$ for all $p$ where 
$\text{Dir}(\cdot)$ denotes the Dirichlet distribution and $\mathbf{1}_{113}$ is a $1\times 113$ vector of ones. Note that for a symmetric Dirichlet distribution with a concentration parameter of 1 across $M$ categories, its density function simplifies to 
\begin{equation}
\frac{1}{C}\prod_{i=1}^M x_i^{1 - 1} = \frac{1}{C},
\end{equation}
where $C$ is a constant. Therefore, assuming $\boldsymbol\alpha_p \sim \text{Dir}(\mathbf{1}_{113})$ is equivalent to applying a uniform prior to $\boldsymbol\alpha_p$, and it follows that it is non-informative. Further, if we let $f(\boldsymbol\alpha_p \mid \mathbf{Z}_p)$ represent the posterior distribution for $\alpha_p$ and $n_{pa}$ be the number of people who experienced health outcome $p$ at age $a$, then we can use the likelihood in \eqref{likelihood} to show that the Dirichlet prior described for $\boldsymbol{\alpha}_p$ is conjugate:  
\begin{align}
f(\boldsymbol{\alpha}_p \mid \mathbf{Z}_p) &\propto \text{f}(\boldsymbol{\alpha}_p)f(\mathbf{Z}_p \mid \boldsymbol\Theta) \notag \\
&\propto \left[\prod_{a=0}^{113}\left(\alpha_p(a)\right)^{1-1}\right]\left[\prod_{a=0}^{113} \left(\alpha_p(a)\right)^{n_{pa}}\right]  \label{apost} \\
&= \prod_{a=0}^{113} \left(\alpha_p(a)\right)^{(n_{pa}+1)-1}. \notag
\end{align} 
In a similar vein, we assume $$\boldsymbol\gamma_p = (\gamma_p(\text{female}), \gamma_p(\text{male}))' \sim \text{Dir}(\mathbf{1}_2)$$ and $$\boldsymbol\xi_p = (\xi_p(\text{Asian}), \dots, \xi_p(\text{White}))' \sim \text{Dir}(\mathbf{1}_4),$$ which, by the same argument, are also non-informative and conjugate priors.

%%% NOTE: STILL NEED TO DESCRIBE WHY IG PRIORS ARE APPROPRIATE %%%%
Our approach for the remaining unknown parameters is to assume vague priors (and conjugate 
where possible). We assume $\{\sigma^2_p\}$ and $\sigma^2_\mu$ are independent 
$\mathcal{IG}\left(2, 1\right)$ 
random variables where $\mathcal{IG}(a,b)$ denotes the inverse-gamma distribution with scale 
$a$ and rate $b$. We also assume that $\boldsymbol{\beta} \sim 
\mathcal{N}(0, 50\mathbf{I}_{Q})$ leading to a Gaussian complete conditional for $\boldsymbol{\beta}$.  
Conjugate prior distributions for each spatial smoothness parameter (the $\nu$'s) and decay 
parameter (the $\psi$'s) are not available; however, following the results of \cite{Zhang2004}, 
a vague prior for the variance parameters (the $\sigma^2$'s) allows each smoothness and decay 
parameter to be fixed without loss generality or predictive power.

\subsection{Model Fitting}
We fit this model within a Bayesian framework. Because the priors
for $\boldsymbol{\alpha}_p$, $\boldsymbol{\gamma}_p$, and $\boldsymbol{\xi}_p$ are all conjugate, their posterior
distributions are known and can be drawn from directly. The posterior densities for the rest of the parameters are unknown and so we obtain draws from the posterior using a Metropolis within Gibbs sampler of the following form:
\begin{enumerate}
  \item \label{draws2} Sample each $\sigma^2_p \sim [ \sigma^2_p | \boldsymbol{\phi}^\star_p, \mathbf{M}_p]$;
  \item \label{drawphi} Sample each $\phi^\star_p \sim [ \phi^\star_p | \sigma^2_p, \mathbf{M}_p ]$;
  \item \label{drawsmu} Sample $\sigma^2_\mu \sim [ \sigma^2_\mu | \boldsymbol{\mu}, \boldsymbol{\beta}, 
    \mathbf{M}_\mu ]$;
  \item \label{drawmu} Sample $\boldsymbol{\mu} \sim [ \boldsymbol{\mu} | \{\boldsymbol{\phi}^\star_p\}, 
    \{\sigma^2_p\}, \{\mathbf{M}_p\}, \sigma^2_\mu, \mathbf{M}_\mu, \boldsymbol{\beta} ]$; and
  \item \label{drawbeta} Sample $\boldsymbol{\beta} \sim [ \boldsymbol{\beta} | \boldsymbol{\mu}, 
    \sigma^2_\mu, \mathbf{M}_\mu ]$.
\end{enumerate}

>>>>

\textcolor{red}{A potential solution to avoid realigning the census data to the grid cells is to let the census block groups equal the $\mathcal{G}_k$ denoted in \eqref{lambda} rather than the grid cells. However, the heat information used in this research (described in detail in Section \ref{sec:temp_data}) is provided at a 1-km$^2$ resolution and these grid cells were chosen to align exactly with the available heat data. Using the block groups in place of the grid cells would necessitate aligning the heat information to the census block group areas. Because heat is our primary variable of interest we prefer realigning the demographic information to the grid associated with the heat data as opposed to the alternative. Therefore, we proceed having aligned the population data as described.}

@@

\subsection{Hyperprior Layer}
\textcolor{red}{At the hyperior prior layer of the model, we outline our prior assumptions for the model
parameters in Sections \ref{likelayer} and \ref{process}. Because $\boldsymbol{\alpha}_p = 
(\alpha_p(0),\dots,\alpha_p(113))'$ are probabilities where $\sum_a \alpha_p(a) = 1$, 
we assume $$\boldsymbol{\alpha}_{p} \sim \text{Dir}(\mathbf{1}_{113})$$ for all $p$ where 
$\text{Dir}(\cdot)$ denotes the Dirichlet distribution and $\mathbf{1}_{113}$ is a $1\times 113$ vector of ones. Note that for a symmetric Dirichlet distribution with a concentration parameter of 1 across $M$ categories, its density function simplifies to 
\begin{equation}
\frac{1}{C}\prod_{i=1}^M x_i^{1 - 1} = \frac{1}{C},
\end{equation}
where $C$ is a constant. Therefore, assuming $\boldsymbol\alpha_p \sim \text{Dir}(\mathbf{1}_{113})$ is equivalent to applying a uniform prior to $\boldsymbol\alpha_p$, and it follows that it is non-informative. Further, if we let $f(\boldsymbol\alpha_p)$ and $f(\boldsymbol\alpha_p \mid \mathbf{Z}_p)$ represent the prior and posterior distributions for $\alpha_p$, respectively, and let $n_{p,a}$ be the number of people who experienced health outcome $p$ at age $a$, then we can use the likelihood in \eqref{likelihood} to show that the Dirichlet prior described for $\boldsymbol{\alpha}_p$ is conjugate:  
\begin{align}
f(\boldsymbol{\alpha}_p \mid \mathbf{Z}_p) &\propto \text{f}(\boldsymbol{\alpha}_p)f(\mathbf{Z}_p \mid \boldsymbol\Theta) \notag \\
&\propto \left[\prod_{a=0}^{113}\left(\alpha_p(a)\right)^{1-1}\right]\left[\prod_{a=0}^{113} \left(\alpha_p(a)\right)^{n_{p,a}}\right]  \label{apost} \\
&= \prod_{a=0}^{113} \left(\alpha_p(a)\right)^{(n_{p,a}+1)-1}. \notag
\end{align} 
Stated explicitly, the posterior distribution of $\boldsymbol\alpha_p$ is a Dirichlet distribution with concentration parameters $(n_{p,0} + 1, n_{p,1} + 1, \dots, n_{p,113}+1)$.
In a similar vein, we assume $$\boldsymbol\gamma_p = (\gamma_p(\text{female}), \gamma_p(\text{male}))' \sim \text{Dir}(\mathbf{1}_2)$$ and $$\boldsymbol\xi_p = (\xi_p(\text{Asian}), \dots, \xi_p(\text{White}))' \sim \text{Dir}(\mathbf{1}_4),$$ which, by the same argument, are also non-informative and conjugate priors.}

%%% NOTE: STILL NEED TO DESCRIBE WHY IG PRIORS ARE APPROPRIATE %%%%
\textcolor{red}{Our approach for the remaining unknown parameters is to assume vague priors (and conjugate 
where possible). This is justified because there is sufficient data to learn these parameters without the need to rely heavily on prior information. We assume $\{\sigma^2_p\}$ and $\sigma^2_\mu$ are independent 
$\mathcal{IG}\left(2.001, 1.001\right)$ 
random variables where $\mathcal{IG}(a,b)$ denotes the inverse-gamma distribution with scale 
$a$ and rate $b$, a well-known conjugate prior for normal variance parameters. The $\mathcal{IG}(2.001, 1.001)$ distribution has a variance of 1000, providing sufficient flexibility for the data to accurately dictate the variances. For the regression coefficients we assume that $\boldsymbol{\beta} \sim 
\mathcal{N}(0, 50\mathbf{I}_{Q})$, leading to a Gaussian complete conditional for $\boldsymbol{\beta}$. 
Conjugate prior distributions for each spatial smoothness parameter (the $\nu$'s) and decay 
parameter (the $\psi$'s) are not available; however, following the results of \cite{Zhang2004}, 
a vague prior for the variance parameters (the $\sigma^2$'s) allows each smoothness and decay 
parameter to be fixed without loss generality or predictive power.}

\subsection{Model Fitting}
\textcolor{red}{We fit this model within a Bayesian framework. Because the priors
for $\boldsymbol{\alpha}_p$, $\boldsymbol{\gamma}_p$, and $\boldsymbol{\xi}_p$ are all conjugate, their posterior
distributions are known and can be drawn from directly. That is, we can easily obtain random draws from $\text{Dir}(n_{p,0} + 1, n_{p,1} + 1, \dots, n_{p,113})$, $\text{Dir}(n_{p,Female} + 1, n_{p,Male} + 1)$, and $\text{Dir}(n_{p,Asian} + 1, \dots, n_{p,White} + 1)$. The posterior densities for the rest of the parameters are unknown and so we obtain draws from the posterior using a Metropolis within Gibbs sampler of the following form:
\begin{enumerate}
  \item \label{draws2} Sample each $\sigma^2_p \sim [ \sigma^2_p | \boldsymbol{\phi}^\star_p, \mathbf{M}_p]$;
  \item \label{drawphi} Sample each $\phi^\star_p \sim [ \phi^\star_p | \sigma^2_p, \mathbf{M}_p ]$;
  \item \label{drawsmu} Sample $\sigma^2_\mu \sim [ \sigma^2_\mu | \boldsymbol{\mu}, \boldsymbol{\beta}, 
    \mathbf{M}_\mu ]$;
  \item \label{drawmu} Sample $\boldsymbol{\mu} \sim [ \boldsymbol{\mu} | \{\boldsymbol{\phi}^\star_p\}, 
    \{\sigma^2_p\}, \{\mathbf{M}_p\}, \sigma^2_\mu, \mathbf{M}_\mu, \boldsymbol{\beta} ]$; and
  \item \label{drawbeta} Sample $\boldsymbol{\beta} \sim [ \boldsymbol{\beta} | \boldsymbol{\mu}, 
    \sigma^2_\mu, \mathbf{M}_\mu ]$.
\end{enumerate}}
